Global Climate of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2017-10-17 ms.tex
caption

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03"/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were 35k photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup.

@@

\section{Light Curve Fits}
We fit a two-component model to the light curves. One component models the astrophysical signal (the planet's thermal phase variation and transit), and the other component models the systematic noise introduced by changes in instrument sensitivity over time. For each light curve, we fit the physical and systematic components simultaneously, such that the total observed flux as a function of time is given by $F(t) = F_\mathrm{physical}(t) \times F_\mathrm{sys}(t)$. For the \HST\ data, where we observed two light curves, we constrain the physical parameters to be the same for both, but allow some of the systematics parameters to vary (see \S\,\ref{sec:hst_sys}).

@@

We explored several different models for the planet's thermal phase variation $P(t)$, using climate maps from the open-source Python package \texttt{SPIDERMAN}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding thermal phase curve over a specified wavelength range. We tested three different maps:  a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). These models each have different strengths. The kinematic model incorporates the most prior knowledge of physics, and even though it makes several simplifying assumptions (e.g, that the wind speed and direction are constant), it reproduces GCM simulations fairly well.   The spherical harmonics map is the most flexible and makes the fewest assumptions about the true climate. The two-temperature model is the simplest and has the fewest free parameters.  In addition to these maps, we also fit the phase variation with a sinusoid, which has been used for many past analyses of exoplanet phase curves (FIXME).
%We explored adding an additional sine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC).

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase.}
\label{fig:ellipsoidal}
\end{figure}

@@

\subsubsection{\HST\ Systematics}
\label{sec:hst_sys}
We fit the WFC3 systematics using an analytic model of the form:
\begin{equation}
 F_\mathrm{sys}(t) = (c\,S(t) + v_1\,t_\mathrm{v} + v_2\,t_\mathrm{v}^2)(1 - \exp(-a\,t_\mathrm{orb} - b))
\end{equation}
where $t_\mathrm{v}$ is time elapsed since the first exposure in a visit and $t_\mathrm{orb}$ is time since the first exposure in an orbit. $S(t)$ is a scale factor equal to 1 for exposures with spatial scanning in the forward direction and $s$ for reverse scans. The orbit-long ramp parameters are consistent for all the visits, so we constrained $a$, $b$, and $s$ to have the same value for all visits in the final fit. The visit-long trends are not consistent, so $c$, $v_1$, and $v_2$ were allowed to vary between visits. We fixed $v_2$ to zero for the two secondary eclipse observations from Program 13360, since the visit-long trend for shorter observations is fit well by a linear slope.

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/systematics.pdf}
\caption{Raw light curves.}
\label{fig:systematics}
\end{figure}

\section{Results}
We achieved

>>>>

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03 arcsec/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were $35\times10^3$ photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup.

@@

\section{Light Curve Fits}
We fit a two-component model to the light curves. One component models the astrophysical signal (the planet's thermal phase variation and transit), and the other component models the systematic noise introduced by time-dependent changes in instrument performance. For each light curve, we fit the physical and systematic components simultaneously, such that the total observed flux as a function of time is given by $F(t) = F_\mathrm{physical}(t) \times F_\mathrm{sys}(t)$. For the \HST\ data, where we observed two phase curves and two additional eclipses, we constrain the physical parameters to be the same for all visits, but allow some of the systematics parameters to vary (for more details see \S\,\ref{sec:hst_sys}). We fit the WFC3 band-integrated (``white" light curve"), as well as the spectroscopic light curves (FIXME describe).

@@

We explored several different models for the planet's thermal phase variation $P(t)$, using climate maps from the open-source Python package \texttt{SPIDERMAN}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding thermal phase curve over a specified wavelength range. We tested three different maps:  a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). These models each have different strengths. The kinematic model incorporates the most prior knowledge of physics, and even though it makes several simplifying assumptions (e.g, that the wind speed and direction are constant), it closely reproduces GCM simulations over a wide parameter space \citep{zhang16}.   The spherical harmonics map is the most flexible and makes the fewest assumptions about the true climate. The two-temperature model is the simplest and has the fewest free parameters.  In addition to these maps, we also fit the phase variation with a sinusoid, which has been used for many past analyses of exoplanet phase curves (FIXME).
%We explored adding an additional sine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC).

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase, normalized to unity at phase zero. The area variation was predicted analytically using the model from \citep{leconte11}.}
\label{fig:ellipsoidal}
\end{figure}

@@

\subsubsection{\HST\ Systematics}
\label{sec:hst_sys}
We fit the WFC3 systematics using an analytic model of the form:
\begin{equation}
 F_\mathrm{sys}(t) = (c\,S(t) + v_1\,t_\mathrm{v} + v_2\,t_\mathrm{v}^2)(1 - \exp(-a\,t_\mathrm{orb} - b))
\end{equation}
where $t_\mathrm{v}$ is time elapsed since the first exposure in a visit and $t_\mathrm{orb}$ is time since the first exposure in an orbit. $S(t)$ is a scale factor equal to 1 for exposures with spatial scanning in the forward direction and $s$ for reverse scans, to account for the upstream-downstream effect (FIXME). The orbit-long ramp parameters are consistent for all the visits, so we constrained $a$, $b$, and $s$ to have the same value for all visits in the final fit. The visit-long trends differ from visit to visit, so $c$, $v_1$, and $v_2$ were allowed to vary between visits. We fixed $v_2$ to zero for the two secondary eclipse observations from Program 13360, since the visit-long trend for shorter observations is fit well by a linear slope.

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/systematics.pdf}
\caption{Raw light curves for WASP-103b observed with \HST/WFC3 (top panel) and \Spitzer/IRAC (bottom two panels). The data points are indicated with black dots. The \HST\ data are unbinned, and the Spitzer data are binned in time segments of FIXME minutes with error bars indicating the bin standard deviation. The colored lines show the best fit models, which include the astrophysical signal and instrument systematics. The gray lines indicate the contribution from the instrument systematics alone (which would be observed for a source with constant brightness and no planet). For visual clarity, we corrected the \HST\ data for the upstream-downstream effect, separated the four visits by adding a flux offset, and zoomed in on the phase variation, so the transits are not displayed in the panel.}
\label{fig:systematics}
\end{figure}

\section{Results}
We achieved good quality fits!
